---
output: github_document 
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prismaread

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of prismaread is allowing to easily import PRISMA L1 hyperspectral images and convert them in a easier to use format (ENVI or GeoTiff).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r install, eval=FALSE}
# install.packages("remotes")
remotes::install_github("lbusett/prismaread")
```
## Usage

The function to be used is `convert_prisma`. It takes as input the full path 
of a PRISMA hdf5 image, an output file name and format, and a series of switches
allowing to decide which datasets should be saved. For example the code:

```{r example, eval=FALSE}
library(prismaread)

 in_file  <- "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
 out_file <- "/home/lb/tmp/test/test_1"
 out_format <- "ENVI"
 
 # Save a full image, prioritizing the SWIR spectrometer and save in EVI format
 convert_prisma(in_file    = in_file,
                out_file   = out_file,
                out_format = out_format,
                FULL = TRUE,
                PAN  = TRUE,
                CLD  = TRUE
                )

```

accesses the input file and saves both the VNIR and SWIR cubes, as well as a full hyperspectral cube and the PAN and CLOUD images. See [documentation](reference/convert_prisma.html) of the convert_prisma() function for info on available arguments. 
