---
output: github_document 
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prismaread

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/lbusett/prismaread.svg?branch=master)](https://travis-ci.org/lbusett/prismaread)
<!-- badges: end -->

The goal of prismaread is allowing to easily import PRISMA L1 hyperspectral images and convert them in a easier to use format (ENVI or GeoTiff).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r install, eval=FALSE}
# install.packages("remotes")
remotes::install_github("lbusett/prismaread")
```
## Work on Level 1 Data

The function to be used is `convert_prisma`. It takes as input the full path 
of a PRISMA hdf5 image, an output file name and format, and a series of switches
allowing to decide which datasets should be saved.

In particular, the `FULL` argument allows deciding if a complete VNIR+SWIR cube has to be created
alongside the "single" VNIR and SWIR ones. In that case, the 'join_priority' 
keyword is used to decide if keeping bands from the "VNIR" or the "SWIR" data
cube in the wavelength were they overlap. 

A "ballpark" georeferencing is provided, but it is not to be trusted/used.

For example the following code:

```{r example, eval=FALSE}
library(prismaread)

in_file  <- "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_file <- "/home/lb/tmp/test/test_1"
out_format <- "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_file   = out_file,
               out_format = out_format,
               join_priority = "VNIR", 
               FULL = TRUE,
               PAN  = TRUE,
               CLOUD  = TRUE)

```

accesses the input file and saves both the VNIR and SWIR cubes, as well as a full hyperspectral cube and the PAN and CLOUD images. __See [documentation](reference/convert_prisma.html) of the convert_prisma() function for info on available arguments__. 

The function also saves ancillary data related to wavelengths and fwhms of the different images, and to hour and sun geometry at acquisition in ancillary txt files. 

### Creation of ATCOR files

Starting `v0.02` the function also allows automatic creation of text files required to run 
an ATCOR atmospheric correction. Those files are saved in the `ATCOR` subfolder of the
main output folder. 

In "standard" behaviour, only the three "standard" ATCOR files (`.wvl`, `.dat` and `.cal`) are created, 
with the `.wvl` file containing nominal wavelengths and FWHMs derived from the `cw` and `fwhm` attributes of the _.he5_ file-
The user can however also choose to generate additional ATCOR files, containing data about
wavelengths and FWHMs related to different "columns" of the data cube, as derived 
from the `KDP_AUX/Cw_Vnir_Matrix`, `KDP_AUX/Cw_Swir_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix` HDF layers. This could allow running different atmospheric 
corrections for different columns of the data, potentially allowing compensating
"smile" effects on the retrieved surface reflectances. For example: 

```{r example2, eval=FALSE}
library(prismaread)

in_file  <- "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_file <- "/home/lb/tmp/test/test_1"
out_format <- "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_file   = out_file,
               out_format = out_format,
               join_priority = "VNIR", 
               ATCOR = TRUE, 
               ATCOR_wls = c(200,800), 
               FULL = TRUE,
               PAN  = TRUE,
               CLOUD  = TRUE)

```

## Work on Level 2 Data

The function to be used is still `convert_prisma`, with similar syntax. Only differences are: 

1. The "PAN", "CLOUD", "LC" and "GLINT" arguments are ignored, because they are 
not available in L2D files; 
2. The "ATCOR" and "ATCOR_wls" arguments are ignored, because they are useless in this case; 
3. The output imagery is properly georeferenced, in case level is 2D. Otherwise, 
 the same "ballpark" georeferencing as for L1 data is applied.
