---
output: github_document 
highlight: pygments
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prismaread

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/lbusett/prismaread.svg?branch=master)](https://travis-ci.org/lbusett/prismaread)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![DOI](https://zenodo.org/badge/211091818.svg)](https://zenodo.org/badge/latestdoi/211091818)
<!-- badges: end -->

The goal of `prismaread` is allowing to easily import PRISMA hyperspectral data (http://www.prisma-i.it/index.php/it/) and convert them to a easier to use format (ENVI or GeoTiff).

_`prismaread` is developed and maintained by Lorenzo Busetto and Luigi Ranghetti, 
[Institute of Remote Sensing of Environment](http://www.irea.cnr.it/en/) - National Research Council - Italy (CNR-IREA)_


# Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r install, eval=FALSE}
# install.packages("remotes")
remotes::install_github("lbusett/prismaread")
library(prismaread)
```
# Work on Level 1 Data

The function to be used is `convert_prisma`. It takes as input the full path 
of a PRISMA hdf5 image, an output file name and format, and a series of switches
allowing to decide which datasets should be saved.

In particular, the `FULL` argument allows deciding if a complete VNIR+SWIR cube has to be created
alongside the "single" VNIR and SWIR ones. In that case, the 'join_priority' 
keyword is used to decide if keeping bands from the "VNIR" or the "SWIR" data
cube in the wavelength were they overlap. 

A "ballpark" georeferencing is provided, but it is __not__ to be trusted/used.
(this may change in the near future)

For example the following code:

```{r example, eval=FALSE}
in_file  = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_file = "/home/lb/tmp/test/test_1"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_file   = out_file,
               out_format = out_format,
               join_priority = "VNIR", 
               FULL = TRUE,
               PAN  = TRUE,
               CLOUD  = TRUE)

```

accesses the input file and saves both the VNIR and SWIR cubes, as well as a full hyperspectral cube and the PAN and CLOUD images. __See [documentation](reference/convert_prisma.html) of the convert_prisma() function for info on available arguments__. 

The function also saves ancillary data related to wavelengths and fwhms of the different images, and to hour and sun geometry at acquisition in ancillary txt files. 

### Creation of ATCOR files

Starting `v0.02` the function also allows automatic creation of text files required to run 
an ATCOR atmospheric correction. Those files are saved in the `ATCOR` subfolder of the
main output folder. 

In "standard" behaviour, only the three "standard" ATCOR files (`.wvl`, `.dat` and `.cal`) are created, 
with the `.wvl` file containing nominal wavelengths and FWHMs derived from the `cw` and `fwhm` attributes of the _.he5_ file-
The user can however also choose to generate additional ATCOR files, containing data about
wavelengths and FWHMs related to different "columns" of the data cube, as derived 
from the `KDP_AUX/Cw_Vnir_Matrix`, `KDP_AUX/Cw_Swir_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix` HDF layers. This could allow running different atmospheric corrections for different columns of the data, potentially allowing compensating
"smile" effects on the retrieved surface reflectances. For example: 

```{r example2, eval=FALSE}
in_file  = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_file = "/home/lb/tmp/test/test_1"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_file   = out_file,
               out_format = out_format,
               join_priority = "VNIR", 
               ATCOR = TRUE, 
               ATCOR_wls = c(200,800), 
               FULL = TRUE,
               PAN  = TRUE,
               CLOUD  = TRUE)

```

__IMPORTANT NOTE__

The latter functionality may be only appliable to "HRC" L1 data cubes. We are currently investigating this - proceed with caution!

# Work on Level 2 Data

The function to be used is still `convert_prisma`, with similar syntax. Only differences are: 

1. The "PAN", "CLOUD", "LC" and "GLINT" arguments are ignored, because they are 
not available in L2 files; 
2. The "ATCOR" and "ATCOR_wls" arguments are ignored, because they are useless in this case; 
3. The output imagery is properly georeferenced, __if level is 2D__. Otherwise, 
the same "ballpark" georeferencing as for L1 data is applied.

# Output Formats

- Outputs are provided as rasters in __ENVI__ or __GEOTIFF__ format according to user's choice. 

- Filenames are built starting from the output file name provided by the user, by
adding appropriate suffixes. For example, if the user specified `out_file = "D:/myoutfolder/myoutfil"`, and `source = "HCO"`, the output file for the VNIR cube will be `D:/myoutfolder/myoutfil_HCO_VNIR.envi` (or
`D:/myoutfolder/myoutfil_HCO_VNIR.tif`).

- Measure units of the output hyperspectral data are as follows: 

  ```{r echo=FALSE, message=FALSE, warning=FALSE}
  my_tbl <- tibble::tribble(
    ~LEVEL, ~Variable, ~"Measure Units",
    "L1",   "Radiance", "W / m^2 * sr * um",
    "L2B",  "Radiance", "W / m^2 * sr * um",
    "L2C",  "Reflectance", "unitless (ratio)",
    "L2D",  "Reflectance", "unitless (ratio)"
  )
  knitr::kable(my_tbl, digits = 3, row.names = FALSE, align = "c",
              caption = NULL)
  ```


- If output format is "ENVI", the wavelengths of the different bands for the hyperspectral 
cubes are properly written in the appropriate header (.hdr) file. 

- Irrespective from output format, info about wavelengths and fwhms of the hyperspectral 
cubes are saved in appropriate txt files. For example, if the output file is 
`D:/myoutfolder/myoutfil_HCO_VNIR.envi`, info about the wavelengths is saved in 
`D:/myoutfolder/myoutfil_HCO_VNIR_meta.txt`

- Info about acquisition date and angles is saved in a dedicated txt file. For example, if
output file is `D:/myoutfolder/myoutfil_VNIR.envi`, info about the angles is saved in 
`D:/myoutfolder/myoutfil_HCO_VNIR_meta.txt`

# Future Work

- Test possibility to more properly georeference L1/L2B/L2C data using the curvilinear
grids functionality in package `stars`

- Implement possibility to apply masks based on the ERR_MATRIX cubes

- Clean up code


## Citation
  
To cite `prismaread` please use:

L. Busetto (2020) prismaread: An R package for imporing PRISMA L1/L2 hyperspectral data and convert them to a more user friendly format, URL: https://github.com/lbusett/prismaread, doi: https://doi.org/10.5281/zenodo.3727671

## Website

For more information, documentation and examples of use, __see also the prismaread website at [https://github.com/lbusett/prismaread/](https://github.com/lbusett/prismaread/)__ 
