---
title: "Computing Spectral Indexes"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(prismaread)
```

`prismaread` allows to automatically compute spectral indexes starting from either an original PRISMA hdf image, or from an hyperspectral cube already processed with `convert prisma`.

## Computing spectral indexes from a predefined list 

Spectral Indexes to be computed can be selected from a list of predefined ones. To do so, either specify a vector of desired indexes as the `indexes` argument of `pr_convert` to compute them directly from the hdf prisma data:

```{r index_1, eval=FALSE, message=FALSE, warning=FALSE}
in_file <- "D:/prismaread/L2D/PRS_L2D_STD_20190616102249_20190616102253_0001.he5"
pr_convert(in_file,
               out_format = "ENVI",
               out_folder = "D:/prismaread/L2D/testL2D/ENVI", 
               indexes = c("GI", "MTCI"))
```

, or use function `pr_compute_indexes` to compute them starting from an hyperspectral cube already processed:

```{r index_2, eval=FALSE}
in_file  <- "PRS_L2D_STD_20190616102249_20190616102253_0001_HCO_FULL.envi"
out_file <- "D:/prismaread/L2D/testL2D/ENVI/outfile_indexes"
pr_compute_indexes(in_file,
                       out_file = out_file,
                       out_format = "ENVI",
                       indexes = c("GI", "MTCI"))
```

Output file names are created by adding a suffix corresponding to the indexes names to the base output filename. In addition, an ancillary file containing the formulas used to compute each index is saved (extension = *.formulas*) for reference. the file shows the formulas used, giving reference to the true wavelengths of the PRISMA bands used in the computation.

The list of available indexes, shown below, was derived from the list of spectral indexes available in package [`hsdar`](https://cran.r-project.org/web/packages/hsdar/hsdar.pdf), as well as from the list reported [HERE](https://cubert-gmbh.com/applications/vegetation-indices/).

```{r tbl, echo=FALSE, message=FALSE, warning=FALSE}

pr_listindexes()

```

**NOTE** Although a check was done on indexes formulas, we do not guarantee that all of them are correct. Please check the formulas in the table to be sure!

**IMPORTANT NOTE**

Computation of spectral indexes can be done alongside extraction of spectral cubes/ancillary info. For example, this commands: 

```{r index2, eval=FALSE}
in_file  <- "PRS_L2D_STD_20190616102249_20190616102253_0001_HCO_FULL.envi"
out_file <- "D:/prismaread/L2D/testL2D/ENVI/outfile_indexes"
pr_compute_indexes(in_file,
                       out_file = out_file,
                       VNIR = TRUE, LATLON = TRUE, ANGLES = TRUE,
                       out_format = "ENVI",
                       indexes = c("GI", "MTCI"))
```

will save in output the VNIR cube, LATLON and ANGLES datasets and compute and save
images relative to the "GI" and "MTCI" indexes. 

### Adding a new index to the list

Additional spectral indexes can be added to the aforementioned list using function `pr_addindex()`, as in the following example

```{r l1example4, eval=FALSE}
pr_addindex(Name = "myindex", Formula = "R600 / R700", 
            Description = "My custom Index", 
            Reference = "Me (2020)"
```

## Computing custom spectral indexes

Custom spectral indexes can also be computed on the fly by specifying the `cust_indexes`
argument, as a __named list__, such as in:

```{r l1example2, eval=FALSE}
pr_convert(in_file,
               out_format = "ENVI",
               out_folder = "D:/prismaread/L2D/testL2D/ENVI", 
               indexes = c("GI", "MTCI"), 
               cust_indexes = list(myindex1 = "R500 / R600",
                                   myindex2 = "(R800 - R680) / (R800 + R680)"))
```
