---
title: "Extracting Data Over Vector Features"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`prismaread` provides a very efficient function based on package `exactextractr`(<https://github.com/isciences/exactextractr>) for extracting and summarizing data from the converted hyperspectral cubes over features of vector spatial files. The function allows to compute several statistics, as well as extracting all pixel values, and to save them to RData, CSV or EXCEL files (**See [documentation](reference/pr_extract_spectra.html) of the prisma\_extract\_spectra() function for info on available arguments**).

For example, starting from a VNIR Cube obtained with `pr_convert` and a vector polygon file:

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(prismaread)
library(ggplot2)
library(tmap)
in_file  <- system.file("testdata/prismaread_test_HCO_FULL.tif",
                        package = "prismaread")
in_vect <- system.file("extdata/testdata/polys_prisma.gpkg",
                       package = "prismaread")

# extract base statistics, in "long" format
test <- pr_extract_spectra(in_file, in_vect, id_field = "field_id")
test
# plot results using ggplot
ggplot(test, aes(x = wvl, y = mean)) +
  geom_line(aes(color = ID, group = ID)) +
  facet_wrap(~ID) +
  theme_light()

```

The `stats_format` defines the format used for statistics' output. If "long" (default, as above), 
the output has one column for the ID of the feature, and one column for each statistic. 
If "wide", the output has one column for each ID/statistic couple (e.g., mean_id_1, stdev_id_1, mean_id_2,
etcetera).

```{r echo=TRUE, message=FALSE, warning=FALSE}
# extract base statistics for VNIR and save results as excel file, in "wide" format

in_file  <- system.file("testdata/prisma_test_HCO_VNIR.tif",
                        package = "prismaread")
test <- pr_extract_spectra(in_file, in_vect,
                           out_file = tempfile(fileext = ".xlsx"),
                           stats_format = "wide", id_field = "field_id")
test

```

Additional arguments allow specifying which statistics are needed, as well as 
selecting if info regarding single pixels contained in each polygon need to be 
returned:


```{r echo=TRUE, message=FALSE, warning=FALSE}
# extract custom statistics
test <- pr_extract_spectra(in_file, 
                           in_vect,
                           selstats = c("mean", "coeffvar", "stdev", "min", "max"),
                           id_field = "field_id")

test

# plot results using ggplot
ggplot(test, aes(x = wvl)) +
  geom_line(aes(y = mean, color = ID, group = ID)) +
  geom_line(aes(y = mean + stdev, group = ID), color = "grey75") +
  geom_line(aes(y = mean - stdev, group = ID), color = "grey75") +
  facet_wrap(~ID) +
  theme_light()

# extract custom statistics and quantiles
test <- pr_extract_spectra(in_file, in_vect, quantiles = TRUE,
                           selstats = c("mean", "stdev"),
                           id_field = "field_id")
test


# extract also all pixels
test <- pr_extract_spectra(in_file, in_vect, allpix = TRUE,
                           selstats = c("mean", "stdev"),
                           id_field = "field_id")


# stats are saved in the "stats" slot of the output
test$stats

# pixel values are saved in the "allpix" slot of the output
test$allpix

ggplot(test$allpix, aes(x = wvl)) +
  geom_line(aes(y = value, group = pixel, color = ID), lwd = 0.01)  +
  facet_wrap(~ID) +
  theme_light()

```
