---
title: "Extracting Data Over Vector Features"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`prismaread` provides a very efficient function based on package `exactextractr`(<https://github.com/isciences/exactextractr>) for extracting and summarizing data from the converted hyperspectral cubes over features of vector spatial files. The function allows to compute several statistics, as well as extracting all pixel values, and to save them to RData, CSV or EXCEL files (**See [documentation](reference/pr_extract_spectra.html) of the prisma\_extract\_spectra() function for info on available arguments**).

For example, starting from a VNIR Cube obtained with `pr_convert` and a vector polygon file:

```{r message=FALSE, warning=FALSE}
  library(prismaread)
  library(ggplot2)
  library(tmap)
  in_file <- "D:/prismaread/L2D/testL2D_HCO_VNIR.envi"
  in_vect <- "D:/prismaread/test/testpoints_l2d_polys.gpkg"

  # extract base statistics, in "long" format
  test <- pr_extract_spectra(in_file, in_vect, id_field = "ID")
  test
  # plot results using ggplot
  ggplot(test, aes(x = wvl, y = mean)) +
    geom_line(aes(color = ID, group = ID)) +
    facet_wrap(~ID) +
    theme_light()

  # extract base statistics ands save results as excel file, in "wide" format
  test <- pr_extract_spectra(in_file, in_vect, out_file = "D:/Temp/test1.xlsx",
                                 stats_format = "wide", id_field = "id")
  test

  # extract custom statistics
  test <- pr_extract_spectra(in_file, in_vect,
                                 selstats = c("mean", "coeffvar", "stdev", "min", "max"),
                                 id_field = "id")
  
  test
  
  # plot results using ggplot
  ggplot(test, aes(x = wvl)) +
    geom_line(aes(y = mean, color = ID, group = ID)) +
    geom_line(aes(y = mean + stdev, group = ID), color = "grey75") +
    geom_line(aes(y = mean - stdev, group = ID), color = "grey75") +
    facet_wrap(~ID) +
    theme_light()

  
  # extract custom statistics and quantiles
  test <- pr_extract_spectra(in_file, in_vect, quantiles = TRUE,
                                 selstats = c("mean", "stdev"),
                                 id_field = "id")
  test

  
  # extract also all pixels
  test <- pr_extract_spectra(in_file, in_vect, allpix = TRUE,
                                 selstats = c("mean", "stdev"),
                                 id_field = "id")
  
  
  # stats are saved in the "stats" slot of the output
  test$stats
  
  # pixel values are saved in the "allpix" slot of the output
  test$allpix

  ggplot(test$allpix, aes(x = wvl)) +
    geom_line(aes(y = value, group = pixel, color = ID), lwd = 0.01)  +
    facet_wrap(~ID) +
    theme_light()

```
