---
title: "Importing Level 1 Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The function to be used to import PRISMA L1 data is `convert_prisma`. 
It takes as input the full path of a PRISMA L1 hdf5 image, an output folder name and format, and a series of switches
allowing to decide which hyperspectral cubes and ancillary datasets should be crated.

In particular: 

 - the `VNIR` and `SWIR` logical arguments allow to decide if importing the VNIR and SWIR hyperspectral cubes; 
 - the `FULL` logical argument `FULL` argument allows deciding if a complete VNIR+SWIR cube has to be created
alongside the "single" VNIR and SWIR ones. In that case, the 'join_priority' 
keyword is used to decide if keeping bands from the "VNIR" or the "SWIR" data
cube in the wavelength were they overlap. 
 - the `PAN`, `LATLON`, `CLOUD`, `GLINT` and `LC` logical argument allow to decide which of the 
 corresponding ancillary datasets should be created (see the PRISMA manual for additional info)

A "base" georeferincing in Lat/Lon WGS-84 based on the "GLT and Bowtie Correction" technique used in ENVI, and
described in https://www.harrisgeospatial.com/docs/backgroundgltbowtiecorrection.html, unless the `base_georef`
argument is set to FALSE.

For example the following code:

```{r l1example, eval=FALSE}
in_file    = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_folder = "/home/lb/tmp/test/"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_folder = out_folder,
               out_format = out_format,
               join_priority = "SWIR",
               FULL   = TRUE,
               LATLON = TRUE,
               PAN    = TRUE,
               CLOUD  = TRUE)

```

accesses the input file and saves both the VNIR and SWIR cubes, as well as a full hyperspectral cube and the PAN, ANGLES and CLOUD datasets __See [documentation](reference/convert_prisma.html) of the convert_prisma() function for info on available arguments__. 

The function also saves ancillary data related to wavelengths and fwhms of the different images, and to hour and sun geometry at acquisition in ancillary txt files (See [Output format and Naming Conventions](articles/Output-Format-and-Naming.html). 

## Associating acquisition angles with L1 data

PRIMA L1 data unfortunately does not contain information concerning acquisition angles, that is instead available
for all L2 datasets. However, if both the L1 and any L2 dataset area available `prismaread` allows to associate the
ANGLES data retrieved from the L2 dataset to the L1 one. To do that, the user has to specify the additional
`in_L2_file` in the call to `convert_prisma`, such as in: 

```{r l1example4, eval=FALSE}
in_file    = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
in_L2_file = "/home/lb/tmp/test/PRS_L2B_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_folder = "/home/lb/tmp/test/"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_folder = out_folder,
               out_format = out_format,
               join_priority = "SWIR",
               FULL   = TRUE,
               LATLON = TRUE,
               PAN    = TRUE,
               CLOUD  = TRUE, 
               ANGLES = TRUE)
```

__Note that in this case also the georeferencing information used for the GLT georeferencing is taken 
from the L2 dataset!__

## Importing only selected bands

The `selbands_vnir` e `selbands_swir` arguments allow to select only a specified subset of 
PRISMA bands, by specifying an array of required wavelengths. For example: 

``` {r l1example2, eval=FALSE}
convert_prisma(in_file    = in_file,
               out_folder = out_folder,
               out_format = out_format,
               VNIR = TRUE, 
               SwiR = TRUE, 
               selbands_vnir = c(450,550,650), 
               selbands_swir = c(1000,1330), 
               join_priority = "SWIR",
               FULL   = TRUE)
```

will create a 3-band VNIR cube, a 2-band SWIR and a 5-band FULL dataset, by selecting the original 
PRISMA bands whit wavelengths closer to the requested ones. 


## Creation of ATCOR files

When working on L1 data, the `convert_prisma` function also allows automatic creation of text files required to run 
an atmospheric correction using ATCOR. Those files are saved in the "ATCOR" subfolder of the
main output folder. 

In "standard" behaviour, only the three "standard" ATCOR files (`.wvl`, `.dat` and `.cal`) are created,
within the "ATCOR" subfolder of the main output folder, with the `.wvl` file containing nominal wavelengths and FWHMs derived from the `cw` and `fwhm` attributes of the _.he5_ file. 

For example, this code: 

```{r atcor1, eval=FALSE}
in_file  = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_folder = "/home/lb/tmp/test/"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_file   = out_folder,
               out_format = out_format,
               join_priority = "VNIR", 
               ATCOR = TRUE, 
               FULL = TRUE,
               PAN  = TRUE,
               CLOUD  = TRUE)

```

will create input files for ATCOR useful for correction of the full hyperspectral cube. 

The user can however also choose to generate additional ATCOR files, containing data about
wavelengths and FWHMs related to different "columns" of the data cube, as derived 
from the `KDP_AUX/Cw_Vnir_Matrix`, `KDP_AUX/Cw_Swir_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix` HDF layers. This could allow running different atmospheric corrections for different columns of the data, potentially allowing compensating
"smile" effects on the retrieved surface reflectances. For example, this code: 

```{r atcor2, eval=FALSE}
in_file  = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_file = "/home/lb/tmp/test/test_1"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_file   = out_file,
               out_format = out_format,
               join_priority = "VNIR", 
               ATCOR = TRUE, 
               ATCOR_wls = c(200,800), 
               FULL = TRUE,
               PAN  = TRUE,
               CLOUD  = TRUE)

```

will create additional ATCOR imput files with wavelengths corresponding to those of the columns 200 and 800.


__IMPORTANT NOTE__

The latter functionality may be only appliable to "HRC" L1 data cubes. We are currently investigating this - proceed with caution!
