---
title: "Importing Level 1 Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(prismaread)
```

The function to be used to import __PRISMA L1 data__ is `pr_convert`. 

It takes as input the full path of a PRISMA L1 hdf5 image, an output folder 
name and format, and a series of switches which allow deciding which hyperspectral 
cubes and ancillary datasets should be crated.

In particular: 
- the `VNIR` and `SWIR` logical arguments allow deciding if importing the VNIR 
    and SWIR hyperspectral cubes; 
- the `FULL` logical argument allows deciding if a complete VNIR+SWIR cube has 
    to be created.

In that case (`FULL = TRUE`):
- the `join_priority` keyword is used to decide if keeping bands from the "VNIR" 
    or the "SWIR" data cube in the wavelengths' region where they overlap;
- the `PAN`, `LATLON`, `CLOUD`, `GLINT`, `LC` and `ERR_MATRIX` logical arguments 
    allow deciding which of the corresponding ancillary datasets should be created 
    (see the PRISMA manual for additional info);
- the `apply_errmatric` logical argument allows deciding if pixels for which 
    the `ERR_MATRIX` values are above one are to be set to NA; 
- the `base_georef` logical argument allows deciding if a "base" georeferencing 
    in Lat/Lon WGS-84 based on the "GLT and Bowtie  Correction" technique used 
    in ENVI and described [here](https://www.harrisgeospatial.com/docs/backgroundgltbowtiecorrection.html) 
    is applied (if set to `FALSE`, the original 1000 x 1000 datasets are 
    returned -- flipped to orient them "north/south" -- without projection.

All logical arguments are set to `FALSE` by default (with the exception of 
`base_georef`), allowing complete customisation of a `pr_convert` run.

For example, the following code accesses the input L1 file and saves the VNIR 
and SWIR cubes and the PAN, ANGLES and CLOUD datasets.
__See [`pr_convert()` documentation](reference/pr_convert.html) for info on 
available arguments__. 

__IMPORTANT NOTE:__ to run this, you'd need to download the example data from 
GitHub.
The data is about 1 GB, so it could take some time.
The data would be placed in subfolder _testdata_ of the `{prismaread}` 
installation folder.

```{r l1example, echo=TRUE, message=FALSE, warning=FALSE}

l1_he5_path <- file.path(
  system.file("testdata", package = "prismaread"),
  "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001.he5"
)
# The example file is downloaded from Github if not already downloaded. 
# WARNING: this may need a long time (1 GB).
if (!file.exists(l1_he5_path)) {
  message("Downloading test data - this may need a long time...")
  # Install package piggyback if missing - for downloaing GitHub attachments
  if (!requireNamespace("piggyback", quietly = TRUE)) {
    install.packages("piggyback")
  }
  l1_zip_path <- file.path(
    system.file("testdata", package = "prismaread"),
    "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001.zip"
  )
  piggyback::pb_download(
    "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001.zip",
    repo = "irea-cnr-mi/prismaread",
    dest = dirname(l1_zip_path)
  )
  unzip(l1_zip_path, exdir = dirname(l1_he5_path))
  unlink(l1_zip_path)
}
l1_out_dir <- file.path(tempdir(), "prismaread/L1")
dir.create(dirname(l1_out_dir), showWarnings = FALSE)

pr_convert(
  in_file = l1_he5_path,
  out_folder = l1_out_dir,
  out_format = "GTiff",
  VNIR = TRUE,
  SWIR = TRUE,
  LATLON = TRUE,
  PAN = TRUE,
  CLOUD = TRUE
)
```

Output files are saved in `l1_out_dir` according to conventions described 
[here](articles/Output-Format-and-Naming.Rmd) and can then be accessed, 
visualised and processed using standard `R` syntax (e.g., with `raster` or 
`stars`).

```{r, echo=TRUE, message=FALSE, warning=FALSE}
list.files(l1_out_dir)
l1_vnir_path <- file.path(
  l1_out_dir, 
  "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.tif"
)
l1_vnir <- raster::brick(l1_vnir_path)
l1_vnir

mapview::viewRGB(l1_vnir, 40,30,20)
```

The function also exports in text files ancillary data related to:
1. wavelengths and fwhms of the different images;
2. hour and sun geometry at acquisition. 

```{r}
l1_vnir_wvl <- read.table(
  file.path(
    l1_out_dir, 
    "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.wvl"
  ), 
  header = TRUE
)
DT::datatable(l1_vnir_wvl)
```



## Associating acquisition angles with L1 data

PRISMA L1 data unfortunately does not contain information concerning 
acquisition angles, that is instead available for all L2 datasets. 
However, if both the L1 and any L2 dataset area available, `{prismaread}` 
allows associating the ANGLES data retrieved from the L2 dataset to the L1 one. 

To do that, the user has to specify the additional argument 
`in_L2_file` in the call to `pr_convert()`. 
__Note that, in this case, also the georeferencing information used for the 
GLT georeferencing is taken from the L2 dataset.__ 

```{r l1example2, echo=TRUE, message=FALSE, warning=FALSE}
l2_he5_path <- file.path(
  system.file("testdata", package = "prismaread"),
  "PRS_L2C_STD_20200524103704_20200524103708_0001.he5"
)
# The example file is downloaded from Github if not already downloaded. 
# WARNING: this may need a long time (1 GB).
if (!file.exists(l2_he5_path)) {
  message("Downloading test data - this may need a long time...")
  l2_zip_path <- file.path(
    system.file("testdata", package = "prismaread"),
    "PRS_L2C_STD_20200524103704_20200524103708_0001.zip"
  )
  piggyback::pb_download(
    "PRS_L2C_STD_20200524103704_20200524103708_0001.zip",
    repo = "irea-cnr-mi/prismaread",
    dest = dirname(l2_zip_path)
  )
  unzip(l2_zip_path, exdir = dirname(l2_zip_path))
  unlink(l2_zip_path)
}
l12_out_dir <- file.path(tempdir(), "prismaread/L12")

# Save only VNIR, also including the ANGLES dataset, in ENVI format, 
# taking ANGLES and georeferencing info from a L2 file. 
pr_convert(
  in_file = l1_he5_path,
  in_L2_file = l2_he5_path,
  out_folder = l12_out_dir,
  out_format = "ENVI",
  VNIR = TRUE,
  ANGLES = TRUE
)
```


## Importing only selected bands

Arguments `selbands_vnir` and `selbands_swir` allow selecting a specified 
subset of PRISMA bands, by specifying an array of required wavelengths. 
For example, the code below creates a 3-band VNIR cube, a 2-band SWIR and a 
5-band FULL dataset, by selecting the original PRISMA bands with wavelengths 
closer to the requested ones. 

```{r l1example3, echo=TRUE, message=FALSE, warning=FALSE}
l1sel_out_dir <- file.path(tempdir(), "prismaread/L1sel")
pr_convert(
  in_file = l1_he5_path,
  out_folder = l1sel_out_dir,
  out_format = "GTiff",
  VNIR = TRUE, 
  SWIR = TRUE,
  FULL = TRUE,
  selbands_vnir = c(450,550,650,750), 
  selbands_swir = c(1000,1330)
)

list.files(l1sel_out_dir)
l1sel_vnir <- raster::brick(
  file.path(
    l1sel_out_dir, 
    "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.tif"
  )
)
l1sel_vnir

raster::plotRGB(l1sel_vnir, 4,3,2, stretch = "lin")

l1sel_vnir_wvl <- read.table(
  file.path(
    l1sel_out_dir, 
    "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.wvl"
  ), 
  header = TRUE
)
DT::datatable(l1sel_vnir_wvl)
```



## Creation of ATCOR files

When working on L1 data, function `pr_convert()` also allows automatic creation 
of text files required to run an atmospheric correction using ATCOR. 
Those files are saved in the "ATCOR" subfolder of the main output folder. 

In default behaviour, only the three "standard" ATCOR files (`.wvl`, `.dat` and
`.cal`) are created in the "ATCOR" subfolder of the main output folder, 
with the `.wvl` file containing nominal wavelengths and FWHMs derived from the 
`cw` and `fwhm` attributes of the `.he5` file. 

For example, the code below creates input files for ATCOR useful for correction 
of the full hyperspectral cube, and places them in the "ATCOR" subfolder of the 
main output folder. 

```{r atcor1, eval=FALSE}
# Save a full image, prioritising the VNIR spectrometer and saving in ENVI format
pr_convert(
  in_file = l1_he5_path,
  out_folder = l1_out_dir,
  out_format = "GTiff",
  join_priority = "VNIR", 
  ATCOR = TRUE, 
  FULL = TRUE,
  PAN = TRUE,
  CLOUD = TRUE
)
```


The user can also choose to generate additional ATCOR files, containing data 
about wavelengths and FWHMs related to different columns of the data cube, 
as derived from the `KDP_AUX/Cw_Vnir_Matrix`, `KDP_AUX/Cw_Swir_Matrix`,
`KDP_AUX/Cw_Fwhm_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix` HDF layers. 
This could allow running different atmospheric corrections for different columns 
of the data, potentially allowing compensating "smile" effects on the retrieved 
surface reflectances. 

For example, the code below creates additional ATCOR input files with 
wavelengths corresponding to those of the columns 200 and 800.

```{r atcor2, eval=FALSE}
l1sel2_out_dir <- file.path(tempdir(), "prismaread/L1sel2")
# Save a full image, prioritising the VNIR spectrometer and saving in ENVI format
pr_convert(
  in_file = l1_he5_path,
  out_folder = l1sel2_out_dir,
  out_format = "ENVI",
  join_priority = "SWIR", 
  ATCOR = TRUE, 
  ATCOR_wls = c(200,800), 
  FULL = TRUE,
  PAN = TRUE, 
  CLOUD = TRUE
)

```

__IMPORTANT NOTE:__ the latter functionality is only appliable to "HRC" L1 
data cubes.
