---
title: "Importing Level 2 Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The function to be used to import PRISMA L2(B,C or D) data is `convert_prisma`. 
It takes as input the full path of a PRISMA L2 hdf5 image, an output folder name and format, and a series of switches
allowing to decide which hyperspectral cubes and ancillary datasets should be crated.

In particular: 

 - the `VNIR` and `SWIR` logical arguments allow to decide if importing the VNIR and SWIR hyperspectral cubes; 
 - the `FULL` logical argument `FULL` argument allows deciding if a complete VNIR+SWIR cube has to be created
alongside the "single" VNIR and SWIR ones. In that case, the 'join_priority' 
keyword is used to decide if keeping bands from the "VNIR" or the "SWIR" data
cube in the wavelength were they overlap. 
 - the `PAN`, `LATLON` and `ANGLES` allow to decide which of the 
 corresponding ancillary datasets should be created (see the PRISMA manual for additional info)

If working with __L2B__ or __L2C__ data, a "base" georeferincing in Lat/Lon WGS-84 based on the "GLT and Bowtie Correction" technique used in ENVI, and
described in https://www.harrisgeospatial.com/docs/backgroundgltbowtiecorrection.html, unless the `base_georef`
argument is set to FALSE.

If working with __L2D__, the output datasets are already georeferenced (usually in UTM projection), although 
accuracy of geolocation should be checked. 

For example the following code:

```{r l2example, eval=FALSE}
in_file    = "/home/lb/tmp/test/PRS_L2B_STD_20190825103112_20190825103117_0001.he5"
out_folder = "/home/lb/tmp/test/"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_folder = out_folder,
               out_format = out_format,
               VNIR   = TRUE, 
               SWIR   = FALSE,
               FULL   = FALSE,
               LATLON = TRUE,
               PAN    = TRUE,
               ANGLES  = TRUE)

```

The following code accesses the input file and saves the VNIR and SWIR cubes, as well as a full hyperspectral cube and the ANGLES and LATLON datasets __See [documentation](reference/convert_prisma.html) of the convert_prisma() function for info on available arguments__.

```{r l2example2, eval=FALSE}
in_file    = "/home/lb/tmp/test/PRS_L2D_STD_20190825103112_20190825103117_0001.he5"
out_folder = "/home/lb/tmp/test/"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_folder = out_folder,
               out_format = out_format,
               VNIR   = TRUE, 
               SWIR   = TRUE,
               FULL   = TRUE,
               LATLON = TRUE,
               ANGLES  = TRUE)

```

The function also saves ancillary data related to wavelengths and fwhms of the different images, 
and to hour and sun geometry at acquisition in ancillary txt files (See [Output format and Naming Conventions](articles/Output-Format-and-Naming.html). 

## Importing only selected bands

The `selbands_vnir` e `selbands_swir` arguments allow to select only a specified subset of 
PRISMA bands, by specifying an array of required wavelengths. For example: 

``` {r l1example2, eval=FALSE}
convert_prisma(in_file    = in_file,
               out_folder = out_folder,
               out_format = out_format,
               VNIR = TRUE, 
               SwiR = TRUE, 
               selbands_vnir = c(450,550,650), 
               selbands_swir = c(1000,1330), 
               join_priority = "SWIR",
               FULL   = TRUE)
```

will create a 3-band VNIR cube, a 2-band SWIR and a 5-band FULL dataset, by selecting the original 
PRISMA bands whit wavelengths closer to the requested ones. 
