---
title: "Importing Level 1 Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(prismaread)
```

The function to be used to import __PRISMA L1 data__ is `pr_convert`. 

It takes as input the full path of a PRISMA L1 hdf5 image, an output folder name and format, and a series of switches
allowing to decide which hyperspectral cubes and ancillary datasets should be crated.

In particular: 

- the `VNIR` and `SWIR` logical arguments allow to decide if importing the VNIR and SWIR hyperspectral cubes; 
- the `FULL` logical argument allows deciding if a complete VNIR+SWIR cube has to be created.
In that case, the `join_priority` keyword is used to decide if keeping bands from the "VNIR" or the "SWIR" data
cube in the wavelengths' region were they overlap. If `FULL` is `TRUE` and 
- the `PAN`, `LATLON`, `CLOUD`, `GLINT`, `LC` and `ERR_MATRIX` logical arguments allow to decide which of the 
corresponding ancillary datasets should be created (see the PRISMA manual for additional info)
- the `apply_errmatric` logical argument allows to decide if pixels for which the ERR_MATRIX values are
above one are to be set to NoData. 
- the `base_georef` logical argument allows to decide if a "base" georeferencing in Lat/Lon WGS-84 based on the "GLT and Bowtie  Correction" technique used in ENVI (described in https://www.harrisgeospatial.com/docs/backgroundgltbowtiecorrection.html), is  applied. If set to `FALSE`, the original 1000 x 1000 datasets are returned (flipped to orient them "north/south"), without   projection.

All logical arguments are set to `FALSE` by default (with the exception of `base_georef`), allowing complete customization of a `pr_convert` run.

For example the following code accesses the input L1 file and saves a the VNIR and SWIR cubes and the PAN, ANGLES and CLOUD datasets __See [documentation](reference/pr_convert.html) of the pr_convert() function for info on available arguments__. 

__IMPORTANT NOTE__ _To run this, you'd need to download the example data from GitHub. The 
data is about 1 GB so it could take a while!! The data would be placed in in subfolder
_testdata_ of the `prismaread` installation folder._

```{r l1example, echo=TRUE, message=FALSE, warning=FALSE}

testfile_l1 <- file.path(system.file("testdata/", package = "prismaread"),
                         "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001.he5")
# Here we download the example file from github if not already downloaded. 
# WARNING ! This may need a long time (1GB file)
if (!file.exists(testfile_l1)){
  message("Downloading test data - This may need a long time!")
  piggyback::pb_download("PRS_L1_STD_OFFL_20200524103704_20200524103708_0001.zip",
                         repo = "lbusett/prismaread",
                         dest = file.path(system.file("", package = "prismaread"), "/testdata"))
  zipfile <- file.path(system.file("testdata/", package = "prismaread"),
                       "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001.zip")
  unzip(zipfile, exdir = dirname(testfile_l1))
  unlink(zipfile)
}
out_folder = file.path(tempdir(), "prismaread_l1")

pr_convert(in_file    = testfile_l1,
           out_folder = out_folder,
           out_format = "GTiff",
           VNIR       = TRUE,
           SWIR       = TRUE,
           LATLON     = TRUE,
           PAN        = TRUE,
           CLOUD      = TRUE)

```

Output files are saved in `out_folder` according to conventions described [here](articles/Output-Format-and-Naming.Rmd) and can then be accessed, visualized and processed using standard `R` syntax (e.g., with `raster` or `stars`).

```{r}

list.files(out_folder)
vnir  <- raster::brick(file.path(out_folder, "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.tif"))
vnir

mapview::viewRGB(vnir, 40,30,20)
```

The function also saves ancillary data related to wavelengths and fwhms of the different images, and to hour and sun geometry at acquisition in ancillary txt files . 

```{r}
wvls <- read.table(file.path(
  out_folder, 
  "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.wvl"), 
  header = TRUE)
DT::datatable(wvls)
```



## Associating acquisition angles with L1 data

PRISMA L1 data unfortunately does not contain information concerning acquisition angles, that is instead available
for all L2 datasets. However, if both the L1 and any L2 dataset area available `prismaread` allows to associate the
ANGLES data retrieved from the L2 dataset to the L1 one. To do that, the user has to specify the additional
`in_L2_file` in the call to `pr_convert`. 
__Note that in this case also the georeferencing information used for the GLT georeferencing is taken 
from the L2 dataset!__ 

```{r l1example2, echo=TRUE, message=FALSE, warning=FALSE}
testfile_l2 <- file.path(system.file("testdata/", package = "prismaread"),
                         "PRS_L2C_STD_20200524103704_20200524103708_0001.he5")
# Here we download the example file from github if not already downloaded. 
# WARNING ! This may need a long time (1GB file)
if (!file.exists(testfile_l2)){
  message("Downloading test data - This may need a long time!")
  piggyback::pb_download("PRS_L2C_STD_20200524103704_20200524103708_0001.zip",
                         repo = "lbusett/prismaread",
                         dest = file.path(system.file("", package = "prismaread"), "/testdata"))
  zipfile <- file.path(system.file("testdata/", package = "prismaread"),
                       "PRS_L2C_STD_20200524103704_20200524103708_0001.zip")
  unzip(zipfile, exdir = dirname(testfile_l1))
  unlink(zipfile)
}

out_folder = file.path(tempdir(), "prismaread_l12")

# Save a only VNIR,  also including the ANGLES dataset, in ENVI format, taking
# ANGLES and georeferencing info from a L2 file. 
pr_convert(in_file    = testfile_l1,
           in_L2_file = testfile_l2, 
           out_folder = out_folder,
           out_format = "ENVI",
           VNIR   = TRUE,
           ANGLES = TRUE)
```


## Importing only selected bands

The `selbands_vnir` e `selbands_swir` arguments allow to select only a specified subset of 
PRISMA bands, by specifying an array of required wavelengths. For example: 

```{r l1example3, echo=TRUE, message=FALSE, warning=FALSE}
out_folder = file.path(tempdir(), "prismaread_l1sel")
pr_convert(in_file    = testfile_l1,
           out_folder = out_folder,
           out_format = "GTiff",
           VNIR = TRUE, 
           SWIR = TRUE,
           FULL = TRUE,
           selbands_vnir = c(450,550,650,750), 
           selbands_swir = c(1000,1330))
```

will create a 3-band VNIR cube, a 2-band SWIR and a 5-band FULL dataset, by selecting the original 
PRISMA bands whit wavelengths closer to the requested ones. 

```{r}
list.files(out_folder)
r <- raster::stack(file.path(out_folder, "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.tif"))
r

raster::plotRGB(r, 4,3,2, stretch = "lin")

DT::datatable(read.table(file.path(
  out_folder, 
  "PRS_L1_STD_OFFL_20200524103704_20200524103708_0001_HCO_VNIR.wvl")), 
  header = TRUE)
```



## Creation of ATCOR files

When working on L1 data, the `pr_convert` function also allows automatic creation of text files required to run 
an atmospheric correction using ATCOR. Those files are saved in the "ATCOR" subfolder of the
main output folder. 

In "standard" behaviour, only the three "standard" ATCOR files (`.wvl`, `.dat` and `.cal`) are created,
within the "ATCOR" subfolder of the main output folder, with the `.wvl` file containing nominal wavelengths and FWHMs derived from the `cw` and `fwhm` attributes of the _.he5_ file. 

For example, this code: 

```{r atcor1, eval=FALSE}

# Save a full image, prioritizing the VNIR spectrometer and save in ENVI format
pr_convert(in_file       = testfile_l1,
           out_folder    = out_folder,
           out_format    = "GTiff",
           join_priority = "VNIR", 
           ATCOR         = TRUE, 
           FULL          = TRUE,
           PAN           = TRUE,
           CLOUD         = TRUE)

```

will create input files for ATCOR useful for correction of the full hyperspectral cube and place them in the
"ATCOR" subfolder of the main output folder. 

The user can  also choose to generate additional ATCOR files, containing data about
wavelengths and FWHMs related to different columns of the data cube, as derived 
from the `KDP_AUX/Cw_Vnir_Matrix`, `KDP_AUX/Cw_Swir_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix`, `KDP_AUX/Cw_Fwhm_Matrix` HDF layers. This could allow running different atmospheric corrections for different columns of the data, potentially allowing compensating
"smile" effects on the retrieved surface reflectances. For example, this code: 

```{r atcor2, eval=FALSE}
in_file  = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_folder = "/home/lb/pro"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in ENVI format
pr_convert(in_file    = in_file,
           out_folder   = out_folder,
           out_format = out_format,
           join_priority = "SWIR", 
           ATCOR = TRUE, 
           ATCOR_wls = c(200,800), 
           FULL = TRUE,
           PAN  = TRUE, 
           CLOUD  = TRUE)

```

will create additional ATCOR input files with wavelengths corresponding to those of the columns 200 and 800.


__IMPORTANT NOTE__

The latter functionality is only appliable to "HRC" L1 data cubes.
