---
title: "Importing Level 2 Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(prismaread)
```

The function to be used to import __PRISMA L2 data__ (2B, 2C, 2D) is `pr_convert`. 

It takes as input the full path of a PRISMA L1 hdf5 image, an output folder name and format, and a series of switches
allowing to decide which hyperspectral cubes and ancillary datasets should be crated.

In particular: 

- the `VNIR` and `SWIR` logical arguments allow deciding if importing the VNIR and SWIR hyperspectral cubes; 
- the `FULL` logical argument allows deciding if a complete VNIR+SWIR cube has to be created.

In that case (`FULL = TRUE`):
- the `join_priority` keyword is used to decide if keeping bands from the "VNIR" 
    or the "SWIR" data cube in the wavelengths' region where they overlap;
- the `PAN`, `LATLON` and `ANGLES` logical arguments 
    allow deciding which of the corresponding ancillary datasets should be created 
    (see the PRISMA manual for additional info);
- if working with __L2B__ or __L2C__ data,
    the `base_georef` logical argument allows deciding if a "base" georeferencing 
    in Lat/Lon WGS-84 based on the "GLT and Bowtie  Correction" technique used 
    in ENVI and described [here](https://www.harrisgeospatial.com/docs/backgroundgltbowtiecorrection.html) 
    is applied (if set to `FALSE`, the original 1000 x 1000 datasets are 
    returned -- flipped to orient them "north/south" -- without projection.
    If working with __L2D__, the output datasets are already georeferenced 
    (UTM projection), although __accuracy of geolocation should be checked__. 

For example the following code accesses the input L2D file and saves the VNIR 
and SWIR cubes and the PAN, ANGLES and CLOUD datasets.
__See [documentation](reference/pr_convert.html) of the `pr_convert()` function
for info on available arguments.__

__IMPORTANT NOTE:__ to run this, you'd need to download the example data from 
GitHub.
The data is about 1 GB, so it could take some time.
The data would be placed in subfolder _testdata_ of the `{prismaread}` 
installation folder.

```{r l2example, echo=TRUE, message=FALSE, warning=FALSE}
l2d_he5_path <- file.path(
  system.file("testdata", package = "prismaread"),
  "PRS_L2D_STD_20200524103704_20200524103708_0001.he5"
)

# Download and unzip using piggyback if necessary
if (!file.exists(l2d_he5_path)) {
  message("Downloading test data - this may need a long time...")
  if (!requireNamespace("piggyback", quietly = TRUE)) {
    install.packages("piggyback")
  }
  l2d_zip_path <- file.path(
    system.file("testdata", package = "prismaread"),
    "PRS_L2D_STD_20200524103704_20200524103708_0001.zip"
  )
  piggyback::pb_download(
    "PRS_L2D_STD_20200524103704_20200524103708_0001.zip",
    repo = "irea-cnr-mi/prismaread",
    dest = dirname(l2d_zip_path)
  )
  piggyback::pb_track(glob = "inst/testdata/*.zip, inst/testdata/*.he5")
  unzip(l2d_zip_path, exdir = dirname(l2d_he5_path))
  unlink(l2d_zip_path)
}
l2d_out_dir <- file.path(tempdir(), "prismaread/L2D")
dir.create(dirname(l2d_out_dir), showWarnings = FALSE)

# Save a full image, prioritising the VNIR spectrometer and save in ENVI format
pr_convert(
  in_file = l2d_he5_path,
  out_folder = l2d_out_dir,
  out_format = "GTiff",
  VNIR = TRUE,
  SWIR = TRUE,
  LATLON = TRUE,
  ANGLES = TRUE
)
```

Output files are saved in `out_folder` according to conventions described 
[here](articles/Output-Format-and-Naming.Rmd) and can then be accessed, 
visualised and processed using standard `R` syntax (e.g., with `raster` or `stars`).

```{r echo=TRUE, message=FALSE, warning=FALSE}

list.files(l2d_out_dir)
l2d_vnir_path <- file.path(
  l2d_out_dir,
  "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_VNIR.tif"
)
l2d_vnir <- raster::brick(l2d_vnir_path)
l2d_vnir

mapview::viewRGB(l2d_vnir, 40,30,20)
```

The function also saves in text files ancillary data related to:
1. wavelengths and fwhms of the different images:
2. hour and sun geometry at acquisition. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
l2d_vnir_wvl <- read.table(
  file.path(
    l2d_out_dir, 
    "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_VNIR.wvl"
  ), 
  header = TRUE
)
DT::datatable(l2d_vnir_wvl)
```

The following code accesses the input file and saves the VNIR and SWIR cubes, 
as well as a full hyperspectral cube and the ANGLES and LATLON datasets.
__See [`pr_convert()` documentation](reference/pr_convert.html) for info on 
available arguments__. 

```{r l2example2, eval=FALSE}

# Save a full image, prioritising the SWIR spectrometer (default)
# and saves in ENVI format.
pr_convert(
  in_file = l2d_he5_path,
  out_folder = l2d_out_dir,
  out_format = "ENVI",
  VNIR = TRUE, 
  SWIR = TRUE,
  FULL = TRUE,
  LATLON = TRUE,
  ANGLES = TRUE
)

```

## Importing only selected bands

The `selbands_vnir` and `selbands_swir` arguments allow selecting only a specified subset of 
PRISMA bands, by specifying an array of required wavelengths. For example, the following code
will create a 3-band VNIR cube, a 2-band SWIR and a 5-band FULL dataset, by selecting the
original PRISMA bands with wavelengths closer to the requested ones. 

```{r message=FALSE, warning=FALSE, echo=TRUE}
pr_convert(
  in_file = l2d_he5_path,
  out_folder = l2d_out_dir,
  out_format = "GTiff",
  VNIR = TRUE, 
  SWIR = TRUE, 
  selbands_vnir = c(450,550,650), 
  selbands_swir = c(1000,1330), 
  join_priority = "SWIR",
  FULL = TRUE,
  overwrite = T
)

list.files(l2d_out_dir)
l2dsel_vnir <- raster::brick(
  file.path(
    l2d_out_dir, 
    "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_VNIR.tif"
  )
)
l2dsel_vnir

l2dsel_swir <- raster::stack(
  file.path(
    l2d_out_dir, 
    "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_SWIR.tif"
  )
)
l2dsel_vnir
```
