---
title: "Importing Level 1 Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The function to be used to import PRISMA L1 data is `convert_prisma`. 
It takes as input the full path of a PRISMA L1 hdf5 image, an output folder name and format, and a series of switches
allowing to decide which hyperspectral cubes and ancillary datasets should be crated.

In particular: 

 - the `VNIR` and `SWIR` logical arguments allow to decide if importing the VNIR and SWIR hyperspectral cubes; 
 - the `FULL` logical argument `FULL` argument allows deciding if a complete VNIR+SWIR cube has to be created
alongside the "single" VNIR and SWIR ones. In that case, the 'join_priority' 
keyword is used to decide if keeping bands from the "VNIR" or the "SWIR" data
cube in the wavelength were they overlap. 
 - the `PAN`, `LATLON`, `CLOUD`, `GLINT` and `LC` logical argument allow to decide which of the 
 corresponding ancillary datasets should be created (see the PRISMA manual for additional info)

A "base" georeferincing in Lat/Lon WGS-84 based on the "GLT and Bowtie Correction" technique used in ENVI, and
described in https://www.harrisgeospatial.com/docs/backgroundgltbowtiecorrection.html, unless the `base_georef`
argument is set to FALSE.

For example the following code:

```{r l1example, eval=FALSE}
in_file    = "/home/lb/tmp/test/PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5"
out_folder = "/home/lb/tmp/test/"
out_format = "ENVI"

# Save a full image, prioritizing the VNIR spectrometer and save in EVI format
convert_prisma(in_file    = in_file,
               out_folder = out_folder,
               out_format = out_format,
               join_priority = "SWIR",
               FULL   = TRUE,
               LATLON = TRUE,
               PAN    = TRUE,
               CLOUD  = TRUE)

```

accesses the input file and saves both the VNIR and SWIR cubes, as well as a full hyperspectral cube and the PAN, ANGLES and CLOUD datasets __See [documentation](reference/convert_prisma.html) of the convert_prisma() function for info on available arguments__. 

The function also saves ancillary data related to wavelengths and fwhms of the different images, and to hour and sun geometry at acquisition in ancillary txt files. 

### Associating acquisition angles with L1 data

TODO TODO TODO TODO
