---
title: "Importing Level 2 Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(prismaread)
```

The function to be used to import __PRISMA L2 data__ (2B, 2C, 2D) is `pr_convert`. 

It takes as input the full path of a PRISMA L1 hdf5 image, an output folder name and format, and a series of switches
allowing to decide which hyperspectral cubes and ancillary datasets should be crated.

In particular: 

- the `VNIR` and `SWIR` logical arguments allow to decide if importing the VNIR and SWIR hyperspectral cubes; 
- the `FULL` logical argument allows deciding if a complete VNIR+SWIR cube has to be created.
In that case, the `join_priority` keyword is used to decide if keeping bands from the "VNIR" or the "SWIR" data
cube in the wavelengths' region were they overlap. If `FULL` is `TRUE` and 
- the `PAN`, `LATLON` and `ANGLES` allow to decide which of the 
corresponding ancillary datasets should be created (see the PRISMA manual for additional info)
- If working with __L2B__ or __L2C__ data the `base_georef` logical argument allows to decide if a "base" georeferencing in 
Lat/Lon WGS-84 based on the "GLT and Bowtie  Correction" technique used in ENVI (described in  
https://www.harrisgeospatial.com/docs/backgroundgltbowtiecorrection.html), is  applied. If set to `FALSE`, the original 1000 x   1000 datasets are returned (flipped to orient them "north/south"), without   projection. If working with __L2D__, the output 
datasets are already georeferenced (UTM projection), although __accuracy of geolocation should be checked__. 

For example the following code accesses the input L2D file and saves the VNIR and SWIR cubes and the PAN, ANGLES and CLOUD datasets __See [documentation](reference/pr_convert.html) of the pr_convert() function for info on available arguments__. 

__IMPORTANT NOTE__ _To run this, you'd need to download the example data from GitHub. The 
data is about 1 GB so it could take a while!! The data would be placed in in subfolder
_testdata_ of the `prismaread` installation folder._

```{r l2example, echo=TRUE, message=FALSE, warning=FALSE}
testfile_l2D <- file.path(system.file("testdata/", package = "prismaread"),
                          "PRS_L2D_STD_20200524103704_20200524103708_0001.he5")

# Download and unzip using piggyback if necessary
if (!file.exists(testfile_l2D)){
  message("Downloading test data - This may need a long time!")
  piggyback::pb_download("PRS_L2D_STD_20200524103704_20200524103708_0001.zip",
                         repo = "lbusett/prismaread",
                         dest = file.path(system.file("", package = "prismaread"), "/testdata"))
  piggyback::pb_track(glob = "inst/testdata/*.zip, inst/testdata/*.he5")
  zipfile <- file.path(system.file("testdata/", package = "prismaread"),
                       "PRS_L2D_STD_20200524103704_20200524103708_0001.zip")
  unzip(zipfile, exdir = dirname(testfile_l2D))
  unlink(zipfile)
}
out_folder_L2D <- file.path(tempdir(), "prismaread/L2D")
dir.create(out_folder_L2D, recursive = TRUE)

# Save a full image, prioritizing the VNIR spectrometer and save in ENVI format
pr_convert(in_file    = testfile_l2D,
           out_folder = out_folder_L2D,
           out_format = "GTiff",
           VNIR       = TRUE, 
           SWIR       = TRUE,
           LATLON     = TRUE,
           ANGLES     = TRUE)
```

Output files are saved in `out_folder` according to conventions described [here](articles/Output-Format-and-Naming.Rmd) and can then be accessed, visualized and processed using standard `R` syntax (e.g., with `raster` or `stars`).

```{r echo=TRUE, message=FALSE, warning=FALSE}

list.files(out_folder_L2D)
vnir  <- raster::brick(file.path(out_folder_L2D, "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_VNIR.tif"))
vnir

mapview::viewRGB(vnir, 40,30,20)
```

The function also saves ancillary data related to wavelengths and fwhms of the different images, and to hour and sun geometry at acquisition in ancillary txt files . 

```{r echo=TRUE, message=FALSE, warning=FALSE}
wvls <- read.table(file.path(
  out_folder_L2D, 
  "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_VNIR.wvl"))
DT::datatable(wvls)
```

The following code accesses the input file and saves the VNIR and SWIR cubes, as well as a full hyperspectral cube and the ANGLES and LATLON datasets __See [documentation](reference/pr_convert.html) of the pr_convert() function for info on available arguments__.

```{r l2example2, eval=FALSE}

# Save a full image, prioritizing the SWIR spectrometer (default 
# and saves in ENVI format
pr_convert(in_file    = testfile_l2D,
           out_folder = out_folder_L2D,
           out_format = "ENVI",
           VNIR       = TRUE, 
           SWIR       = TRUE,
           FULL       = TRUE,
           LATLON     = TRUE,
           ANGLES     = TRUE)

```

## Importing only selected bands

The `selbands_vnir` and `selbands_swir` arguments allow to select only a specified subset of 
PRISMA bands, by specifying an array of required wavelengths. For example the following code
will create a 3-band VNIR cube, a 2-band SWIR and a 5-band FULL dataset, by selecting the
original PRISMA bands with wavelengths closer to the requested ones. 

```{r message=FALSE, warning=FALSE, echo=TRUE}
pr_convert(in_file       = testfile_l2D,
           out_folder    = out_folder_L2D,
           out_format    = "GTiff",
           VNIR          = TRUE, 
           SWIR          = TRUE, 
           selbands_vnir = c(450,550,650), 
           selbands_swir = c(1000,1330), 
           join_priority = "SWIR",
           FULL          = TRUE,
           overwrite = T)

list.files(out_folder_L2D)
r <- raster::stack(file.path(out_folder_L2D, "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_VNIR.tif"))
r

r <- raster::stack(file.path(out_folder_L2D, "PRS_L2D_STD_20200524103704_20200524103708_0001_HCO_SWIR.tif"))
r

```


