---
title: "Output Format and Naming"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Format of output rasters

Outputs are provided as rasters in __ENVI__ or __GeoTiff__ format according to
user's choice in the `output_format` argument of `pr_convert()`. 

### Naming conventions

Filenames are built starting from the output folder name provided by the user, by
adding appropriate suffixes. If `out_basefile` is not specified (or set to `auto`, the Default), 
output file names are created based on the filename of the hdf5 input. For example, if 
in_file is `PRS_L1_STD_OFFL_20190825103112_20190825103117_0001.he5`, the output file for the VNIR cube will be `PRS_L1_STD_OFFL_20190825103112_20190825103117_0001_HCO_VNIR.envi` (or `.tif`), that for SWIR will be
`PRS_L1_STD_OFFL_20190825103112_20190825103117_0001_HCO_SWIR.envi` ), etcetera. 

If `out_basefile` is specified, the filenames are created substituting the specified string. For example, if 
out_basefile is "myprisma", the output file for the VNIR cube will be `myprisma_HCO_VNIR.envi` (or .tif), that for SWIR will be
`myprisma_HCO_SWIR.envi` ), etcetera.

### Measure Units

- Measure units of the output hyperspectral data are as follows: 

  ```{r echo=FALSE, message=FALSE, warning=FALSE}
  my_tbl <- tibble::tribble(
    ~LEVEL, ~Variable, ~"Measure Units",
    "L1",   "Radiance", "W / m^2 * sr * um",
    "L2B",  "Radiance", "W / m^2 * sr * um",
    "L2C",  "Reflectance", "unitless (ratio)",
    "L2D",  "Reflectance", "unitless (ratio)"
  )
  knitr::kable(my_tbl, digits = 3, row.names = FALSE, align = "c",
              caption = NULL)
  ```

### Ancillary info

- If output format is "ENVI", the wavelengths of the different bands for the hyperspectral 
cubes are properly written in the appropriate header (.hdr) file. 

- Irrespective from output format, info about wavelengths and fwhms of the hyperspectral 
cubes are saved in appropriate txt files. For example, if the output file is 
`D:/myoutfolder/myoutfil_HCO_VNIR.envi`, info about the wavelengths is saved in 
`D:/myoutfolder/myoutfil_HCO_VNIR.wvl`

- Info about acquisition date and angles is saved in a dedicated txt file. For example, if
output file is `D:/myoutfolder/myoutfil_VNIR.envi`, info about the angles is saved in 
`D:/myoutfolder/myoutfil_HCO_VNIR.ang`
